Generalized Bradley-Terry models

### 2025-02-18 inst/jss_hyper3_real.Rnw
work on the diagram

\node at (4.8, 3) {$a$};
\node at (4.8, 2) {$b$};
\node at (4.8, 1) {$c$};
\node at (4.8, 0) {$d$};
\node at (4.8,-1) {$e$};

@@

\fill  (4.8, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill  (4.8, 2) circle[radius=2pt];  % terminal node
\fill  (4.8, 1) circle[radius=2pt];  % terminal node
\fill  (4.8, 0) circle[radius=2pt];  % d finishes; paths abce
\fill  (4.8,-1) circle[radius=2pt];  % terminal node

\draw (4.8,3) -- (5.8,5);
\draw (4.8,3) -- (5.8,4);
\draw (4.8,3) -- (5.8,3);
\draw (4.8,3) -- (5.8,2);

\node at (4.5,4.6) {$b$};
\node at (4.5,3.7) {$c$};
\node at (4.5,3.2) {$d$};
\node at (4.5,2.7) {$e$};

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (6, 5)  {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (6, 4)  {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (6, 3)  {$\frac{p_d     }{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (6, 2)  {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_d+p_e}$};

\fill (7.4, 5) circle[radius=2pt];  % ab finishes; pahts cde
\draw (7.4,5) -- (8.4,6);
\draw (7.4,5) -- (8.4,5);
\draw (7.4,5) -- (8.4,4);

\node at (7.9,5.7) {$p_c$};
\node at (7.9,5.2) {$p_d$};
\node at (7.9,4.7) {$p_e$};
\node at (9.3, 6) {$\frac{\la{p_c}}{\la{p_c}+p_d+p_e}$};
\node at (9.3, 5) {$\frac{p_d  }{\la{p_c}+p_d+p_e}$};
\node at (9.3, 4) {$\frac{p_e  }{\la{p_c}+p_d+p_e}$};

\fill (10.2, 6) circle[radius=2pt];  % abc finishes; paths de
\draw (10.2, 6) -- (11.2,7);
\node at (10.7,6.8) {$p_d$};

\node at (11.8, 7) {$\frac{p_d}{p_d+p_e}$};

\draw (10.2, 6) -- (11.2,6.2);
\node at (11.8, 6.2) {$\frac{p_e}{p_d+p_e}$};
\node at (10.7,6.3) {$p_e$};

\fill (10.2, 5) circle[radius=2pt];
\draw (10.2, 5) -- (11.2,5.3);
\node at (11.8, 5.3) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};
\node at (10.7,5.3) {$p_c$};

\draw (10.2, 5) -- (11.2,4.7);
\node at (11.8, 4.7) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};
\node at (10.7,4.7) {$p_e$};

\fill (10.2, 4) circle[radius=2pt];
\draw (10.2, 4) -- (11.2,3.9);
\node at (11.8, 3.9) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};
\node at (10.7,4.1) {$p_c$};

\draw (10.2, 4) -- (11.2,3.1);
\node at (11.8, 3.1) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};
\node at (10.7,3.7) {$p_e$};

\draw (4, 0) -- (5,1);
\node at (6, 1)  {$\frac{p_a}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4, 0) -- (5,0);
\node at (6, 0)  {$\frac{p_b}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4, 0) -- (5,-1);
\node at (6, -1)  {$\frac{p_c}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4, 0) -- (5,-2);
\node at (6, -2)  {$\frac{\ld{p_e}}{p_a+p_b+p_c+\ld{p_e}}$};

\node at (4.5,0.7) {$a$};
\node at (4.5,0.2) {$b$};
\node at (4.5,-0.3) {$c$};
\node at (4.5,-0.8) {$e$};

\fill (7.4, -2) circle[radius=2pt];  % de finishes; paths abc
\draw (7.4, -2) -- (8.4,-1);
\draw (7.4, -2) -- (8.4,-2);
\draw (7.4, -2) -- (8.4,-3);

\node at (7.9,-1.2) {$p_a$};
\node at (7.9,-1.8) {$p_b$};
\node at (7.9,-2.3) {$p_c$};

\node at (9.3, -1) {$\frac{p_a}{p_a+p_b+c}$};
\node at (9.3, -2) {$\frac{p_b}{p_a+p_b+c}$};
\node at (9.3, -3) {$\frac{p_c}{p_a+p_b+c}$};

\fill (10.2, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.2, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.2, -3) circle[radius=2pt];  % dec finishes; paths ab

\draw (10.2, -1) -- (11.2,-0);
\node at (10.6,-0.4) {$b$};
\node at (10.6,-0.8) {$c$};

\node at (11.8, 0) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}}$};
\draw (10.2, -1) -- (11.2,-1);

\node at (11.8, -1) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}}$};

\draw (10.2, -2) -- (11.2,-1.6);
\node at (11.8, -1.6) {$\frac{\la{p_a}}{\la{p_a}+\la{p_c}}$};

\node at (10.6,-1.7) {$a$};
\node at (10.6,-2) {$c$};

\draw (10.2, -2) -- (11.2,-2.2);
\node at (11.8, -2.2) {$\frac{\la{p_c}}{\la{p_a}+\la{p_c}}$};

\draw (10.2, -3) -- (11.2,-2.9);
\node at (11.8, -2.9) {$\frac{\la{p_a}}{\la{p_a}+\la{p_b}}$};

\draw (10.2, -3) -- (11.2,-3.4);
\node at (11.8, -3.4) {$\frac{\la{p_b}}{\la{p_a}+\la{p_b}}$};

\node at (10.6,-2.8) {$a$};
\node at (10.6,-3.3) {$b$};

\fill (7.4, 1) circle[radius=2pt];  % da finishes; paths bce

\draw (7.4, 1) -- (8.4,2);
\draw (7.4, 1) -- (8.4,1);
\draw (7.4, 1) -- (8.4,0);

\node at (7.9,1.7) {$b$};
\node at (7.9,1.2) {$c$};
\node at (7.9,0.7) {$e$};

\node at (9.3, 2) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.3, 1) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.3, 0) {$\frac{e       }{\la{p_b}+\la{p_c}+p_e}$};

\fill (10.2, 1) circle[radius=2pt];  % da finishes; paths bce
\draw (10.2, 1) -- (11.2,2);
\node at (11.8, 2) {$\frac{\la{p_b}}{\la{p_b}+\ld{p_e}}$};
\draw (10.2, 1) -- (11.2,1);
\node at (11.8, 1) {$\frac{\ld{p_e}}{\la{p_b}+p_e}$};

>>>>

\node at (4.8, 2.8) {$a$};
\node at (4.8, 1.8) {$b$};
\node at (4.8, 0.8) {$c$};
\node at (4.8,-0.2) {$d$};
\node at (4.8,-1.2) {$e$};

@@

\fill[red] (4.8, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill[red] (4.8, 2) circle[radius=2pt];  % terminal node
\fill[red] (4.8, 1) circle[radius=2pt];  % terminal node
\fill[red] (4.8, 0) circle[radius=2pt];  % d finishes; paths abce
\fill[red] (4.8,-1) circle[radius=2pt];  % terminal node

\draw (4.8,3) -- (5.8,5) -- (8.2,5);
\draw (4.8,3) -- (5.8,4) -- (8.2,4);
\draw (4.8,3) -- (5.8,3) -- (8.2,3);
\draw (4.8,3) -- (5.8,2) -- (8.2,2);

\newcommand{\la}[1]{\lambda{#1}}
\newcommand{\ld}[1]{\mu    {#1}}
\node at (7, 5.4)  {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 4.4)  {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 3.4)  {$\frac{p_d     }{\la{p_b}+\la{p_c}+p_d+p_e}$};
\node at (7, 2.4)  {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_d+p_e}$};

\fill[blue] (8.2,5) circle[radius=2pt];  % ab finishes; paths cde
\fill[blue] (8.2,4) circle[radius=2pt];  % ab finishes; paths cde
\fill[blue] (8.2,3) circle[radius=2pt];  % ab finishes; paths cde
\fill[blue] (8.2,2) circle[radius=2pt];  % ab finishes; paths cde
\draw (8.2,5) -- (8.9,6) -- (10.6,6);
\draw (8.2,5) -- (8.9,5) -- (10.6,5);
\draw (8.2,5) -- (8.9,4) -- (10.6,4);

\node at (9.5, 6.4) {$\frac{\la{p_c}}{\la{p_c}+p_d+p_e}$};
\node at (9.5, 5.4) {$\frac{p_d  }{\la{p_c}+p_d+p_e}$};
\node at (9.5, 4.4) {$\frac{p_e  }{\la{p_c}+p_d+p_e}$};

\fill[green] (10.6, 6) circle[radius=2pt];  % abc finishes; paths de
\draw (10.6, 6) -- (11.2,7) -- (12.2,7);
\fill[red] (12.2,7) circle[radius=2pt];
\node at (11.8, 7.4) {$\frac{p_d}{p_d+p_e}$};

\draw (10.6, 6) -- (11.2,6.2) -- (12.2,6.2);
\fill[red] (12.2,6.2) circle[radius=2pt];
\node at (11.8, 6.6) {$\frac{p_e}{p_d+p_e}$};

\fill[green] (10.6, 5) circle[radius=2pt];
\draw (10.6, 5) -- (11.2,5.3) -- (12.2, 5.3);
\fill[red] (12.2,5.3) circle[radius=2pt];
\node at (11.8, 5.7) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw (10.6, 5) -- (11.2,4.7) -- (12.2, 4.7);
\fill[red] (12.2,4.7) circle[radius=2pt];
\node at (11.8, 5.0) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\fill[green] (10.6, 4) circle[radius=2pt];
\draw (10.6, 4) -- (11.2,3.9) -- (12.2, 3.9);
\fill[red] (12.2,3.9) circle[radius=2pt];
\node at (11.8, 4.3) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw (10.6, 4) -- (11.2,3.1) -- (12.2,3.1);
\fill[red] (12.2,3.1) circle[radius=2pt];
\node at (11.8, 3.5) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,1) -- (8.2,1);
\fill[red] (12.2,1) circle[radius=2pt];

\node at (7, 1.4)  {$\frac{p_a}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,0) -- (8.2,0);
\fill[red] (12.2,0) circle[radius=2pt];
\node at (7, 0.4)  {$\frac{p_b}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,-1) -- (8.2,-1);
\node at (7, -0.6)  {$\frac{p_c}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,-2) -- (8.2,-2);
\node at (7, -1.6)  {$\frac{\ld{p_e}}{p_a+p_b+p_c+\ld{p_e}}$};

\fill[blue] (8.2, -2) circle[radius=2pt];  % de finishes; paths abc
\fill[blue] (8.2, -1) circle[radius=2pt];  % de finishes; paths abc
\fill[blue] (8.2,  0) circle[radius=2pt];  % de finishes; paths abc
\draw (8.2, -2) -- (8.9,-1) -- (10.6,-1);
\draw (8.2, -2) -- (8.9,-2) -- (10.6,-2);
\draw (8.2, -2) -- (8.9,-3) -- (10.6,-3);

\fill[red] (12.2,-1) circle[radius=2pt];

\node at (9.5, -0.6) {$\frac{p_a}{p_a+p_b+p_c}$};
\node at (9.5, -1.6) {$\frac{p_b}{p_a+p_b+p_c}$};
\node at (9.5, -2.6) {$\frac{p_c}{p_a+p_b+p_c}$};

\fill[green] (10.6, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill[green] (10.6, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill[green] (10.6, -3) circle[radius=2pt];  % dec finishes; paths ab

\draw (10.6, -1) -- (11.2,-0) -- (12.2,0);

\node at (11.8, 0.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}}$};
\draw (10.6, -1) -- (11.2,-1) -- (12.2,-1);

\node at (11.8, -0.6) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}}$};

\draw(10.6, -2) -- (11.2,-1.8) -- (12.2,-1.8);
\fill[red] (12.2,-1.8) circle[radius=2pt];

\node at (11.8, -1.4) {$\frac{\la{p_a}}{\la{p_a}+\la{p_c}}$};

\draw (10.6, -2) -- (11.2,-2.6) -- (12.2,-2.6);
\fill[red] (12.2,-2.6) circle[radius=2pt];
\node at (11.8, -2.2) {$\frac{\la{p_c}}{\la{p_a}+\la{p_c}}$};

%\draw (10.6, -3) -- (11.2,-2.9);
%\fill[red] (12.2,-2.9) circle[radius=2pt];
%\node[red] at (11.8, -2.9) {$\frac{\la{p_a}}{\la{p_a}+\la{p_b}}$};

\draw (10.6, -3) -- (11.2,-3.4) -- (12.2, -3.4);
\fill[red] (12.2,-3.4) circle[radius=2pt];
\node at (11.8, -3) {$\frac{\la{p_b}}{\la{p_a}+\la{p_b}}$};

\fill[blue] (8.2, 1) circle[radius=2pt];  % da finishes; paths bce

\draw (8.2,1) -- (8.9,2) -- (10.6,2);
\draw (8.2,1) -- (8.9,1) -- (10.6,1);
\draw (8.2,1) -- (8.9,0) -- (10.6,0);

\node at (9.5, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 1.4) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 0.4) {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_e}$};
\fill[green] (10.6, 2) circle[radius=2pt];  % da finishes; paths bce
\fill[green] (10.6, 1) circle[radius=2pt];  % da finishes; paths bce
\fill[green] (10.6, 0) circle[radius=2pt];  % da finishes; paths bce
\draw (10.6, 1) -- (11.2,2) -- (12.2,2);
\node at (11.8, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\ld{p_e}}$};
\draw (10.6, 1) -- (11.2,1) -- (12.2,1);
\node at (11.8, 1.4) {$\frac{\ld{p_e}}{\la{p_b}+p_e}$};
